---
title       : Tweets Word Cloud Generator
subtitle    : 
author      : Veeresh Taranalli
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Motivation, Approach and User Interface

* A word cloud visualization of a given Twitter user's tweets. 
* Useful for getting a simple idea about the user's areas of interest and preferences.
* Some customization allowed for the Shiny App user to explore. 
* Use the `twitteR` package to get tweets corresponding to a user's timeline. 
* Use the `tm` package to clean up the obtained tweets and extract complete words. 
* Use the `wordcloud` package to plot the corresponding wordcloud thus obtained.
* User Interface consists of the following parameters: 
    - `Twitter UserName` 
    - `Number of Previous Tweets?`
    - `Include Re-tweets?`
    - `Include Replies?`
    - `Color Palette`

---

## An Example using R Code

We create a tweet wordcloud from `hackernews` twitter timeline as follows. 

```{r results='hide', warning=FALSE, message=FALSE}
library(twitteR)
library(tm)
library(wordcloud)
library(RColorBrewer)
load("twitter_tokens")
setup_twitter_oauth(consumer_key, consumer_secret, access_token, access_secret)

user_name <- "hackernews"
number_of_tweets <- 100
exclude_replies <- FALSE
include_retweets <- FALSE
tweets_list <- userTimeline(user_name, n = number_of_tweets, includeRts = include_retweets, 
                            excludeReplies = exclude_replies)
tweets_df <- twListToDF(tweets_list)
# build a corpus, and specify the source to be character vectors 
tweets_Corpus <- Corpus(VectorSource(tweets_df$text))
```

```{r echo=FALSE}
# convert to lower 
tweets_Corpus <- tm_map(tweets_Corpus, content_transformer(tolower))
# remove URLs
removeURL <- function(x) gsub("http[^[:space:]]*", "", x)
tweets_Corpus <- tm_map(tweets_Corpus, content_transformer(removeURL)) 
# remove anything other than english letters or space
removeNumPunct <- function(x) gsub("[^[:alpha:][:space:]]*", "", x)
tweets_Corpus <- tm_map(tweets_Corpus, content_transformer(removeNumPunct))
# remove punctuation
tweets_Corpus <- tm_map(tweets_Corpus, removePunctuation)
# remove numbers
tweets_Corpus <- tm_map(tweets_Corpus, removeNumbers)
# # add extra stop words:
tweetsStopwords <- c(stopwords('english'), "like", "absolutely", "want", 
                         "ive", "im", "go", "seen", "via", "get", "amp", "will") 
# remove stopwords from corpus
tweets_Corpus <- tm_map(tweets_Corpus, removeWords, tweetsStopwords, lazy=TRUE)
# remove extra whitespace
tweets_Corpus <- tm_map(tweets_Corpus, stripWhitespace, lazy=TRUE)
```

```{r echo=FALSE}
# Keep a copy
tweets_CorpusCopy <- tweets_Corpus
# Extract Stem Words
tweets_Corpus <- tm_map(tweets_Corpus, stemDocument)
stemCompletion2 <- function(x, dictionary) {
            x <- unlist(strsplit(as.character(x), " "))
            # Unexpectedly, stemCompletion completes an empty string to
            # a word in dictionary. Remove empty string to avoid above issue. 
            x <- x[x != ""]
            x <- stemCompletion(x, dictionary=dictionary)
            x <- paste(x, sep="", collapse=" ") 
            PlainTextDocument(stripWhitespace(x))
        }
        
tweets_Corpus <- tm_map(tweets_Corpus, content_transformer(stemCompletion2), 
                                dictionary=tweets_CorpusCopy, lazy=TRUE)
tweets_Corpus <- Corpus(VectorSource(tweets_Corpus))
tdm <- TermDocumentMatrix(tweets_Corpus, control = list(wordLengths = c(1, Inf)))
```

---

```{r warning=FALSE, message=FALSE, echo=FALSE}
m <- as.matrix(tdm)
# calculate the frequency of words and sort it by frequency 
word.freq <- sort(rowSums(m), decreasing = T)
# colors
pal <- brewer.pal(9, "BuGn")
pal <- pal[-(1:4)]
```

```{r}
# plot word cloud
wordcloud(words = names(word.freq), freq = word.freq, min.freq = 3,
          random.order = F, colors = pal, scale=c(4,.2))

```

---

## Project Code

* The complete source code `ui.R` and `server.R` files for the Shiny App are available at https://github.com/veeresht/ShinyAppCode

